---
title: "Analyzing ToothGrowth data"
author: "Sergio Martínez Tornell"
date: "22 November 2015"
output: pdf_document
---
## Summary
In this report we analyze the ToothGrowth dataset from package _datasets_ in $R$
we found that the outcome len is strongly related with dose and supp. Moreover, for certain _doses_, there is a significant difference for different _supp_.

```{r include=FALSE}
library(datasets)
library(dplyr)
library(ggplot2)
data <- ToothGrowth
```

## Exploratory analysis 

First, get a summary of the data.
```{r}
summary(data)
str(data)
```

The data has 3 columns:

* **len:** Which is the outcome value.
* **supp:** Which seems to be a factor of the experiment.
* **dose:** Which is a numerical factor which takes only three values: $0.5$, $1$, $2$.

Summarize and plot a histogram of the **len**.

```{r echo=FALSE}
summary(data$len)
hist(data$len, xlab= "len")
```

Check whether supp and dose are related

```{r echo=FALSE}
knitr::kable(table(data$supp, data$dose))
```

It seems that both were factors of the experiment.

Do some qplot:

```{r echo=FALSE}
qplot(data$dose, data$len, color = data$supp)
```

It seems at *doses* of 0.5 and 1.0 each *supp* has a slightly different effect on
the outcome. However at *dose* 2.0 it seems there is no difference between
*supps*. Moreover, it seems to be a correlation between *dose* and *len*
independently of the *supp*.

## Analysis

### Objective of the analysis

We want to validate our hypotheses stated at the end of the previous section:
1. There is a difference in the outcome *len* between different *doses*
2. There is a difference in the outcome *len* at *doses* 0.5 and 1.0 between the
different *supps*.
3. There is **NO** difference in the outcome *len* at *dose* 2.0 between the 
different *supps*.


### Validate the impact of *dose* in *len*

To simplify, we test if there is a significant difference between 0.5 and 2.0
*doses*. Since we can not establish a relationship between the measurements, we
have performed a $t.test$ for not paired populations assuming equal variance for
both populations.

```{r}
results <- t.test(data$len[data$dose == 0.5], data$len[data$dose == 2.0],
                  var.equal = TRUE, paired = FALSE)
results
```

We have found that there is a mean difference of 
$`r results$estimate[1]-results$estimate[2]`$ with a p value of $`r results$p.value`$.
A very small $p$ value suggest that this relationship is strong and it is rarely
related to chance alone. We can confirm our hypotheses: **A bigger dose produce a
bigger _len_**.

### Validate the impact of *supp* in *len* at different *doses*

#### Testing the impact at *dose* $0.5$

```{r}
results <- t.test(data$len[data$dose == 0.5 & data$supp == "VC" ], 
                  data$len[data$dose == 0.5 & data$supp == "OJ" ],
                  var.equal = TRUE, paired = FALSE)
results
```
We have found that there is a mean difference ($VJ - OJ$) of 
$`r results$estimate[1]-results$estimate[2]`$ with a p value of $`r results$p.value`$.
A very small $p$ value suggest that this relationship is strong and it is rarely
related to chance alone. We can confirm our hypotheses: **There is different between _supps_ at $0.5$ _dose_**.

#### Testing the impact at *dose* $1.0$

```{r}
results <- t.test(data$len[data$dose == 1 & data$supp == "VC" ], 
                  data$len[data$dose == 1 & data$supp == "OJ" ],
                  var.equal = TRUE, paired = FALSE)
results
```
We have found that there is a mean difference ($VJ - OJ$) of 
$`r results$estimate[1]-results$estimate[2]`$ with a p value of 
$`r results$p.value`$.
A very small $p$ value suggest that this relationship is strong and it is rarely
related to chance alone. We can confirm our hypotheses: **There is different between _supps_ at $1$ _dose_**.

#### Testing the impact at *dose* $2.0$

```{r}
results <- t.test(data$len[data$dose == 2 & data$supp == "VC" ], 
                  data$len[data$dose == 2 & data$supp == "OJ" ],
                  var.equal = TRUE, paired = FALSE)
results
```
We have found that there is a mean difference ($VJ - OJ$) of 
$`r results$estimate[1]-results$estimate[2]`$ with a p value of 
$`r results$p.value`$.
A big $p$ value suggest that this relationship is weak and it is probably
related to chance alone. We can confirm our stated hypotheses: **There is no different between _supps_ at $2.0$ _dose_**.

## Conclusion

In our test we confirmed that:

1. There is a difference in the outcome *len* between different *doses*
2. There is a difference in the outcome *len* at *doses* 0.5 and 1.0 between the
different *supps*.
3. There is **NO** difference in the outcome *len* at *dose* 2.0 between the 
different *supps*.

The discovered relationships suggest that a better analysis considereng 
co-related variables may be desirable.
